Abundance/diversity GAM

### 2021-04-07 code/ms1_abundance.Rmd
updated analyses and figures ot include mamagement; dropped MRM analysis

### 2.2.2 Call models
```{r}
# BB
# gam_max_bb_abund_00 <- gam(max.bb.abund ~ 
#                              s(bb.sp, bs = "re") + # random intercept for bb.sp
#                              s(bb.sp.year, bs = "re") + # random intercept for bb.sp:year
#                              s(elev.mean, by = bb.sp, k = 8), # smoother for elev by bb.sp.
#                            method = "REML",
#                            family = "tw", # Tweedie distribution performs well
#                            select = TRUE,
#                            data = abund_sum) %>% getViz()
# 
# # FL
# gam_mean_fl_abund_00 <- gam(mean.fl.abund ~
#                               year +
#                               s(bb.sp, bs = "re") +
#                               s(elev.mean, by = bb.sp, k = 8),
#                            method = "REML",
#                            family = "tw",
#                            select = TRUE,
#                            data = abund_sum) %>% getViz()

# BB
gam_max_bb_abund_01 <- gam(max.bb.abund ~ 
                             management +
                             s(bb.sp, bs = "re") + # random intercept for bb.sp
                             s(bb.sp.year, bs = "re") + # random intercept for bb.sp:year
                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by bb.sp.
                           method = "REML",
                           family = "nb",
                           select = TRUE,
                           data = abund_sum) %>% getViz()

@@

#### 2.2.4.1 BB
```{r}
# Plot smooths and year effect on the link scale

## Individual plots
gers_abund <- plot(sm(gam_max_bb_abund_01, 3)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "gers")
hort_abund <- plot(sm(gam_max_bb_abund_01, 4)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "hort")
jone_abund <- plot(sm(gam_max_bb_abund_01, 5)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "jone")
mend_abund <- plot(sm(gam_max_bb_abund_01, 6)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mend")
mont_abund <- plot(sm(gam_max_bb_abund_01, 7)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mont")
muci_abund <- plot(sm(gam_max_bb_abund_01, 8)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "muci")
pasc_abund <- plot(sm(gam_max_bb_abund_01, 9)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pasc")
prat_abund <- plot(sm(gam_max_bb_abund_01, 10)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "prat")
psit_abund <- plot(sm(gam_max_bb_abund_01, 11)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "psit")
pyre_abund <- plot(sm(gam_max_bb_abund_01, 12)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pyre")
soro_abund <- plot(sm(gam_max_bb_abund_01, 13)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "soro")
telu_abund <- plot(sm(gam_max_bb_abund_01, 14)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "telu")
wurf_abund <- plot(sm(gam_max_bb_abund_01, 15)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "wurf")
mgmt_bb_effect <- plot(pterm(gam_max_bb_abund_01, 1)) + l_fitPoints() + l_ciBar() + 
  labs(x = NULL, y = NULL, title = "Management")

@@

# Extract predictions and plot on response scale

gam_max_bb_abund_01_ilink <- family(gam_max_bb_abund_01)$linkinv # get inverse link function

gam_max_bb_abund_01_pred <- predict.gam(gam_max_bb_abund_01, # get model predictions
                                        type = "terms") %>% 
  as.data.frame() %>% # convert to data frame
  as_tibble() %>% # convert to tibble
  dplyr::select(mgmt = 1, bb.sp = 2, bb.sp.year = 3, gers = 4, # rename
         hort = 5, jone = 6, mend = 7, mont = 8, 
         muci = 9, pasc = 10, prat = 11, psit = 12, 
         pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
  mutate(fit = # get cumulative fit
           mgmt + bb.sp + bb.sp.year + gers + 
           hort + jone + mend + mont + 
           muci + pasc + prat + psit + 
           pyre + soro + telu + wurf 
           ) %>%
  dplyr::select(fit) %>%
  mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
  bind_cols(abund_sum)

ggplot(gam_max_bb_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
  geom_point(aes(elev.mean, max.bb.abund, shape = tree_line, color = management), 
             inherit.aes = FALSE, alpha = 0.5, size = 2) + # add original data and map tree line to color
  #geom_line(size = 0.5, alpha = 0.5) +
  facet_wrap(~bb.sp, scales = "free_y") +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = bquote("Peak bumble bee abundance"),
       color = "Tree line",
       shape = "Year",
       linetype = "Year")

ggsave("../output/gam_bb_abund.pdf")
ggsave("../output/gam_bb_abund.png")
```

@@

gers_fl_abund <- plot(sm(gam_mean_fl_abund_01, 2)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "gers")
hort_fl_abund <- plot(sm(gam_mean_fl_abund_01, 3)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "hort")
jone_fl_abund <- plot(sm(gam_mean_fl_abund_01, 4)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "jone")
mend_fl_abund <- plot(sm(gam_mean_fl_abund_01, 5)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mend")
mont_fl_abund <- plot(sm(gam_mean_fl_abund_01, 6)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mont")
muci_fl_abund <- plot(sm(gam_mean_fl_abund_01, 7)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "muci")
pasc_fl_abund <- plot(sm(gam_mean_fl_abund_01, 8)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pasc")
prat_fl_abund <- plot(sm(gam_mean_fl_abund_01, 9)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "prat")
psit_fl_abund <- plot(sm(gam_mean_fl_abund_01, 10)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "psit")
pyre_fl_abund <- plot(sm(gam_mean_fl_abund_01, 11)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pyre")
soro_fl_abund <- plot(sm(gam_mean_fl_abund_01, 12)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "soro")
telu_fl_abund <- plot(sm(gam_mean_fl_abund_01, 13)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "telu")
wurf_fl_abund <- plot(sm(gam_mean_fl_abund_01, 14)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "wurf")
mgmt_fl_effect <- plot(pterm(gam_mean_fl_abund_01, 1)) + l_fitPoints() + l_ciBar() + 
  labs(x = NULL, y = NULL, title = "Management")
year_fl_effect <- plot(pterm(gam_mean_fl_abund_01, 2)) + l_fitPoints() + l_ciBar() + 
  labs(x = NULL, y = NULL, title = "Year")

@@

# Extract predictions and plot on response scale

gam_mean_fl_abund_01_ilink <- family(gam_mean_fl_abund_01)$linkinv # get inverse link function to transform to the response scale

gam_mean_fl_abund_01_pred <- predict.gam(gam_mean_fl_abund_01, # get model predictions
                                        type = "terms") %>% 
  as.data.frame() %>% # convert to data frame
  as_tibble() %>% # convert to tibble
  dplyr::select(mgmt = 1, year = 2, bb.sp = 3, gers = 4, # rename
         hort = 5, jone = 6, mend = 7, mont = 8, 
         muci = 9, pasc = 10, prat = 11, psit = 12, 
         pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
  mutate(fit = # get cumulative fit
           mgmt + year + bb.sp + gers + 
           hort + jone + mend + mont + 
           muci + pasc + prat + psit + 
           pyre + soro + telu + wurf 
           ) %>%
  dplyr::select(fit) %>%
  mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
  bind_cols(na.omit(abund_sum)) # the na.omit() drops the rows where floral surveying data were missing

ggplot(gam_mean_fl_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
  geom_point(aes(elev.mean, mean.fl.abund, color = management), inherit.aes = FALSE, alpha = 0.5) + # add original data and map tree line to color
  #geom_line(size = 0.5, alpha = 0.5) +
  facet_wrap(~bb.sp, scales = "free_y") +
  theme_light() +
  labs(x = "Elevation (m.s.l.)", 
       y = bquote("Peak floral abundance"),
       color = "Tree line",
       shape = "Year",
       linetype = "Year")

ggsave("../output/gam_fl_abund.pdf")
ggsave("../output/gam_fl_abund.png")
```

>>>>

### 2.2.2 Call models
```{r}
# BB
gam_max_bb_abund_01 <- gam(max.bb.abund ~ 
                             management +
                             s(bb.sp, bs = "re") + # random intercept for bb.sp
                             s(bb.sp.year, bs = "re") + # random intercept for bb.sp:year
                             s(elev.mean, by = bb.sp, k = 8), # smoother for elev by bb.sp.
                           method = "REML",
                           family = "nb",
                           select = TRUE,
                           data = abund_sum) %>% getViz()

@@

#### 2.2.4.1 BB
```{r}
## Individual plots
gers_abund <- plot(sm(gam_max_bb_abund_01, 3)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "gers") +
  l_rug(length = unit(0.15, "cm"))
hort_abund <- plot(sm(gam_max_bb_abund_01, 4)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "hort") +
  l_rug(length = unit(0.15, "cm"))
jone_abund <- plot(sm(gam_max_bb_abund_01, 5)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "jone") +
  l_rug(length = unit(0.15, "cm"))
mend_abund <- plot(sm(gam_max_bb_abund_01, 6)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mend") +
  l_rug(length = unit(0.15, "cm"))
mont_abund <- plot(sm(gam_max_bb_abund_01, 7)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mont") +
  l_rug(length = unit(0.15, "cm"))
muci_abund <- plot(sm(gam_max_bb_abund_01, 8)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "muci") +
  l_rug(length = unit(0.15, "cm"))
pasc_abund <- plot(sm(gam_max_bb_abund_01, 9)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pasc") +
  l_rug(length = unit(0.15, "cm"))
prat_abund <- plot(sm(gam_max_bb_abund_01, 10)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "prat") +
  l_rug(length = unit(0.15, "cm"))
psit_abund <- plot(sm(gam_max_bb_abund_01, 11)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "psit") +
  l_rug(length = unit(0.15, "cm"))
pyre_abund <- plot(sm(gam_max_bb_abund_01, 12)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pyre") +
  l_rug(length = unit(0.15, "cm"))
soro_abund <- plot(sm(gam_max_bb_abund_01, 13)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "soro") +
  l_rug(length = unit(0.15, "cm"))
telu_abund <- plot(sm(gam_max_bb_abund_01, 14)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "telu") +
  l_rug(length = unit(0.15, "cm"))
wurf_abund <- plot(sm(gam_max_bb_abund_01, 15)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "wurf") +
  l_rug(length = unit(0.15, "cm"))
mgmt_bb_effect <- plot(pterm(gam_max_bb_abund_01, 1)) + l_fitPoints() + l_ciBar() + 
  labs(x = NULL, y = NULL, title = "Management")

@@

# 
# 
# # Extract predictions and plot on response scale
# 
# gam_max_bb_abund_01_ilink <- family(gam_max_bb_abund_01)$linkinv # get inverse link function
# 
# gam_max_bb_abund_01_pred <- predict.gam(gam_max_bb_abund_01, # get model predictions
#                                         type = "terms") %>% 
#   as.data.frame() %>% # convert to data frame
#   as_tibble() %>% # convert to tibble
#   dplyr::select(mgmt = 1, bb.sp = 2, bb.sp.year = 3, gers = 4, # rename
#          hort = 5, jone = 6, mend = 7, mont = 8, 
#          muci = 9, pasc = 10, prat = 11, psit = 12, 
#          pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
#   mutate(fit = # get cumulative fit
#            mgmt + bb.sp + bb.sp.year + gers + 
#            hort + jone + mend + mont + 
#            muci + pasc + prat + psit + 
#            pyre + soro + telu + wurf 
#            ) %>%
#   dplyr::select(fit) %>%
#   mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
#   bind_cols(abund_sum)
# 
# ggplot(gam_max_bb_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
#   geom_point(aes(elev.mean, max.bb.abund, shape = tree_line, color = management), 
#              inherit.aes = FALSE, alpha = 0.5, size = 2) + # add original data and map tree line to color
#   #geom_line(size = 0.5, alpha = 0.5) +
#   facet_wrap(~bb.sp, scales = "free_y") +
#   theme_light() +
#   labs(x = "Elevation (m.s.l.)", 
#        y = bquote("Peak bumble bee abundance"),
#        color = "Tree line",
#        shape = "Year",
#        linetype = "Year")
# 
# ggsave("../output/gam_bb_abund.pdf")
# ggsave("../output/gam_bb_abund.png")
```

@@

gers_fl_abund <- plot(sm(gam_mean_fl_abund_01, 2)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "gers") +
  l_rug(length = unit(0.15, "cm"))
hort_fl_abund <- plot(sm(gam_mean_fl_abund_01, 3)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "hort") +
  l_rug(length = unit(0.15, "cm"))
jone_fl_abund <- plot(sm(gam_mean_fl_abund_01, 4)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "jone") +
  l_rug(length = unit(0.15, "cm"))
mend_fl_abund <- plot(sm(gam_mean_fl_abund_01, 5)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mend") +
  l_rug(length = unit(0.15, "cm"))
mont_fl_abund <- plot(sm(gam_mean_fl_abund_01, 6)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "mont") +
  l_rug(length = unit(0.15, "cm"))
muci_fl_abund <- plot(sm(gam_mean_fl_abund_01, 7)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "muci") +
  l_rug(length = unit(0.15, "cm"))
pasc_fl_abund <- plot(sm(gam_mean_fl_abund_01, 8)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pasc") +
  l_rug(length = unit(0.15, "cm"))
prat_fl_abund <- plot(sm(gam_mean_fl_abund_01, 9)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "prat") +
  l_rug(length = unit(0.15, "cm"))
psit_fl_abund <- plot(sm(gam_mean_fl_abund_01, 10)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "psit") +
  l_rug(length = unit(0.15, "cm"))
pyre_fl_abund <- plot(sm(gam_mean_fl_abund_01, 11)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "pyre") +
  l_rug(length = unit(0.15, "cm"))
soro_fl_abund <- plot(sm(gam_mean_fl_abund_01, 12)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "soro") +
  l_rug(length = unit(0.15, "cm"))
telu_fl_abund <- plot(sm(gam_mean_fl_abund_01, 13)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "telu") +
  l_rug(length = unit(0.15, "cm"))
wurf_fl_abund <- plot(sm(gam_mean_fl_abund_01, 14)) + l_fitLine() + l_ciLine() + 
  labs(x = NULL, y = NULL, title = "wurf") +
  l_rug(length = unit(0.15, "cm"))
mgmt_fl_effect <- plot(pterm(gam_mean_fl_abund_01, 1)) + l_fitPoints() + l_ciBar() + 
  labs(x = NULL, y = NULL, title = "Management")
year_fl_effect <- plot(pterm(gam_mean_fl_abund_01, 2)) + l_fitPoints() + l_ciBar() + 
  labs(x = NULL, y = NULL, title = "Year")

@@

# 
# 
# # Extract predictions and plot on response scale
# 
# gam_mean_fl_abund_01_ilink <- family(gam_mean_fl_abund_01)$linkinv # get inverse link function to transform to the response scale
# 
# gam_mean_fl_abund_01_pred <- predict.gam(gam_mean_fl_abund_01, # get model predictions
#                                         type = "terms") %>% 
#   as.data.frame() %>% # convert to data frame
#   as_tibble() %>% # convert to tibble
#   dplyr::select(mgmt = 1, year = 2, bb.sp = 3, gers = 4, # rename
#          hort = 5, jone = 6, mend = 7, mont = 8, 
#          muci = 9, pasc = 10, prat = 11, psit = 12, 
#          pyre = 13, soro = 14, telu = 15, wurf = 16) %>%
#   mutate(fit = # get cumulative fit
#            mgmt + year + bb.sp + gers + 
#            hort + jone + mend + mont + 
#            muci + pasc + prat + psit + 
#            pyre + soro + telu + wurf 
#            ) %>%
#   dplyr::select(fit) %>%
#   mutate(fit.trans = gam_max_bb_abund_01_ilink(fit)) %>% # back-transform to response scale
#   bind_cols(na.omit(abund_sum)) # the na.omit() drops the rows where floral surveying data were missing
# 
# ggplot(gam_mean_fl_abund_01_pred, aes(elev.mean, fit.trans, linetype = year)) + # map year to linetype
#   geom_point(aes(elev.mean, mean.fl.abund, color = management), inherit.aes = FALSE, alpha = 0.5) + # add original data and map tree line to color
#   #geom_line(size = 0.5, alpha = 0.5) +
#   facet_wrap(~bb.sp, scales = "free_y") +
#   theme_light() +
#   labs(x = "Elevation (m.s.l.)", 
#        y = bquote("Peak floral abundance"),
#        color = "Tree line",
#        shape = "Year",
#        linetype = "Year")
# 
# ggsave("../output/gam_fl_abund.pdf")
# ggsave("../output/gam_fl_abund.png")
```
